<link href="http://kevinburke.bitbucket.org/markdowncss/markdown.css" rel="stylesheet"></link>
Poverty by District in Israel, mapped
---

## 0. Data preparation

Make fake data of the same style as Nepal estimates. But for many fewer districts. Use
Isreali district names.


Lets load up the data and the "NepalMapUtils" library. This works fine if you do a [setwd](http://www.statmethods.net/interface/workspace.html) to inside the `Poverty` folder within the [NepalMaps](http://github.com/prabhasp/NepalMaps) repository.

```{r echo=T, comment=NA, fig.height=6, fig.width=10, cache=TRUE}
# setwd(...) here
poverty11 <- read.csv("ILEstimates2011.csv")
source("../ILMapUtils.R")
```

We will do one tranformation before proceeding, which is to rename our columns. In the dataset, "FGT(0)" (which loads in R as `FGT.0.` because R doesn't like parentheses) is the _poverty incidence_ metric, defined as proportion of individuals living in that area who are in households with an average per capita expenditure below the poverty line. FGT(1) is the _poverty gap_, which is the average distance below the poverty line, being zero for those individuals above the line, and FGT(2) is _poverty severity_, the squared distance for those below hte line, which gives more weight to the very poor. [(See 2006 report for definitions)](http://cbs.gov.np/wp-content/uploads/2012/Others/SAE%20of%20Poverty,%20Caloric%20Intake%20and%20Malnutrition%20in%20Nepal.pdf).

So lets go ahead and rename our columns to these understandable names:
```{r}
names(poverty11)
names(poverty11) <- c("District", "Population", "PovertyIncidence", "S.E-P.I.", "PovertyGap", 
                    "S.E-P.G." , "PovertySeverity", "S.E-P.S.")
```

And now we will also load the 2001 data, which has been modified to have the exact same column names already:

```{r echo=T, comment=NA, fig.height=6, fig.width=10, cache=TRUE}
poverty2001 <- read.csv("ILEstimates2001.csv")
```

## 1. Poverty Incidence (2011)

 Lets make a quick map of it (note that I haven't paid attention to map projections: these are sketches).
```{r echo=T, comment=NA, fig.height=7., fig.width=12, cache=TRUE}
npchoropleth(poverty11, 'District', 'PovertyIncidence')
```

A second map, coloring those that are above the mean (weighted by population) as blue and those below as red:
```{r echo=T, comment=NA, fig.height=7., fig.width=12, cache=TRUE, message=F}
meanpoverty <- weighted.mean(poverty11$PovertyIncidence, poverty11$Population)
npchoropleth(poverty11, 'District', 'PovertyIncidence') + scale_fill_gradient2(low=muted('blue'), midpoint=meanpoverty, mid='white', high=muted('red'))
```

In the second map, you can really see ...

## 2. Absolute poor (2011)

The next thing to look at, as Chandan did, is the number of absolute poor, which is easily calculable given that population is nicely included in this dataset. Lets have a look:
```{r echo=T, comment=NA, fig.height=7.2, fig.width=12, cache=TRUE}
poverty11$AbsolutePoor <- poverty11$Population * poverty11$PovertyIncidence
npchoropleth(poverty11, 'District', 'AbsolutePoor')
```
The absolute poor are concentrated in ...

For reference, a population sketch to remind us where people live in Israel:
```{r echo=T, comment=NA, fig.height=7.2, fig.width=12, cache=TRUE}
npchoropleth(poverty11, 'District', 'Population')
```

3. Comparisons with 2001
===

Now, lets us compare the data to the 2001 small area povery estimates. We will first create a new dataframe containing data from both years, then take the difference and map it (while remembering to flip the color scheme, a decrease is poverty incidence is good!):

```{r echo=T, comment=NA, fig.height=7.2, fig.width=12, cache=TRUE, message=FALSE}
poverty <- merge(poverty11, poverty2001, by="District", suffixes=c("_2011", "_2001"))
poverty$PovertyTenYrChange <- poverty$PovertyIncidence_2011 - poverty$PovertyIncidence_2001
npchoropleth(poverty, 'District', 'PovertyTenYrChange') + 
  scale_fill_gradient2(low=muted('blue'), high=muted('red'))
```

The units here are in difference of percentage points; bigger negative numbers are better. The message here seems to be that most of the country has become less poor, with an exception of an increase in poverty in: Mustang / Manang, the Far Western Mountains, and in the central-eastern Terai (Parsa/Bara/Rautahat and Siraha/Saptari).
